PANCHANGE_data_preprocessing

### 2020-08-14 PANCHANGE_data_preprocessing2020-07-15.Rmd
fixed NBR IDs

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = FALSE
RAMP = TRUE
```

@@

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(tidyverse)
library(eeptools) #Needed to calculate age from birthdate
```

@@

# GLAD demographics
## Read in data
```{r glad dem cols}
if(GLAD == TRUE){
glad.dem.raw <- readRDS(file = paste0(data.raw_path, "/glad/dem_glad.rds"))
dim(glad.dem.raw)
colnames(glad.dem.raw)
}
```

## Select variables
```{r glad dem cleaning}
if(GLAD == TRUE){
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.prepandemic",
                  "Age_uncleaned",
                  "EduYrs")

@@

# EDGI demographics
## Read in data
```{r edgi dem cols}
if(EDGI == TRUE){
  edgi.dem.raw <- readRDS(file = paste0(data.raw_path, "/edgi/dem_edgi.rds"))
  dim(edgi.dem.raw)
  colnames(edgi.dem.raw)
}
```

@@

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.edgi = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.edgi, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## Select variables
```{r nbr dem cleaning}
if(NBR == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.nbr",
                        "Age_uncleaned",
                        "EduYrs")

@@

```{r nbr cohort cleaning}
if(NBR == TRUE) {
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.coping",
                  "Age_uncleaned",
                  "EduYrs")

nbr.cohort.raw.id <- nbr.cohort.raw %>%
  drop_na(`NBR ID`) %>% # Drop NAs
  distinct(`NBR ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`NBR ID`, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID),
#         startDate.coping = startDate
         ) %>% # Rename startDate
  select(
         Sample, # Sample
         ID, # ID
         Cohort = cohort, #cohort information
#         startDate.coping, #Date when participant signed up to NBR COPING
#         Age_uncleaned = age, # Age
#         Gender = gender, # Gender
#         Sex = demographics.select_questionnaire_items_medical, #Sex
#         EduYrs = demographics.school_education_school_years.txt, # Years in education
#         demographics.college_or_university_degree, # level of education - univeristy
#         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#         demographics.cses_or_equivalent, # level of education - cse
#         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity.unc = ethnicity # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# Select variables
```{r ramp dem cleaning}
if(RAMP == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.ramp")

@@

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.ramp, #Date when participant signed up to ramp
      age_category = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
     # EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
     # demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
     # demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
    #  demographics.o_levelsgcses_or_equivalent, # level of education - gcse
     # demographics.cses_or_equivalent, # level of education - cse
     # demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# Select variables
```{r ramp employment cleaning}
if(RAMP == TRUE){
  exclude_cols_employment <- c("ID",
                        "Sample",
                        "startDate.ramp")

@@

ramp.employment.raw.id <- ramp.employment.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      Sample,
      ID,
      startDate.ramp,
      highest_education_ramp = employment.what_is_your_highest_level_of_education, #this is missing NVQ or equivalent which we have in GLAD-COPING
      # EduYrs = employment.university_education_attend_school.txt, # Years in education - not in RAMP dataset
      employment.government_work_key_workers
    ) %>%
    add_numeric(., exclude = exclude_cols_employment) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

nbr.employment.raw.id <- nbr.employment.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  #mutate(
  #  ID = as.numeric(ID) ##NBR IDs have a letter at the end so can't be coerced to numeric
  #  ) %>% 
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r glad mhd cleaning}
if(GLAD == TRUE) {
glad.mhd.raw.id <- glad.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r edgi mhd cleaning}
if(EDGI == TRUE) {
edgi.mhd.raw.id <- edgi.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

nbr.mhd.raw.id <- nbr.mhd.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(#ID = as.numeric(ID), ## NBR IDs end with letters so can't be numeric
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

>>>>

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = TRUE
NBR = FALSE
RAMP = FALSE
```

@@

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(eeptools) # Needed to calculate age from birthdate
library(tidyverse)
```

@@

# Demographics
## GLAD
### Read in data
```{r glad dem cols}
if(GLAD == TRUE){
glad.dem.raw <- readRDS(file = paste0(data.raw_path, "/glad/dem_glad.rds"))
dim(glad.dem.raw)
colnames(glad.dem.raw)
}
```

### Select variables
```{r glad dem cleaning}
if(GLAD == TRUE){
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.prepandemic",
                  "Age_uncleaned",
                  "EduYrs")

@@

## EDGI
### Read in data
```{r edgi dem cols}
if(EDGI == TRUE){
  edgi.dem.raw <- readRDS(file = paste0(data.raw_path, "/edgi/dem_edgi.rds"))
  dim(edgi.dem.raw)
  colnames(edgi.dem.raw)
}
```

@@

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.prepandemic = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.prepandemic, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## Select variables
```{r nbr dem cleaning}
if(NBR == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.coping",
                        "Age_uncleaned",
                        "EduYrs")

@@

```{r nbr cohort cleaning}
if(NBR == TRUE) {
exclude_cols_cohort_nbr <- c("ID",
                  "Sample",
                  "Age_uncleaned",
                  "EduYrs")

nbr.cohort.raw.id <- nbr.cohort.raw %>%
  drop_na(`NBR ID`) %>% # Drop NAs
  distinct(`NBR ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`NBR ID`, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         Cohort = cohort, #cohort information
#         startDate.coping, #Date when participant signed up to NBR COPING
#         Age_uncleaned = age, # Age
#         Gender = gender, # Gender
#         Sex = demographics.select_questionnaire_items_medical, #Sex
#         EduYrs = demographics.school_education_school_years.txt, # Years in education
#         demographics.college_or_university_degree, # level of education - univeristy
#         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#         demographics.cses_or_equivalent, # level of education - cse
#         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity.unc = ethnicity # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_cohort_nbr) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# Select variables
```{r ramp dem cleaning}
if(RAMP == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.coping")

@@

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.coping = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping, #Date when participant signed up to ramp
      age_category = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
     # EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
     # demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
     # demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
    #  demographics.o_levelsgcses_or_equivalent, # level of education - gcse
     # demographics.cses_or_equivalent, # level of education - cse
     # demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# Select variables
```{r ramp employment cleaning}
if(RAMP == TRUE){
  exclude_cols_employment <- c("ID",
                        "Sample")

@@

ramp.employment.raw.id <- ramp.employment.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample,
      ID,
      highest_education_ramp = employment.what_is_your_highest_level_of_education, #this is missing NVQ or equivalent which we have in GLAD-COPING
      # EduYrs = employment.university_education_attend_school.txt, # Years in education - not in RAMP dataset
      employment.government_work_key_workers
    ) %>%
    add_numeric(., exclude = exclude_cols_employment) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

nbr.employment.raw.id <- nbr.employment.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r glad mhd cleaning}
if(GLAD == TRUE) {
glad.mhd.raw.id <- glad.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r edgi mhd cleaning}
if(EDGI == TRUE) {
edgi.mhd.raw.id <- edgi.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

nbr.mhd.raw.id <- nbr.mhd.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

  separate(recipientFirstName, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(ID = as.numeric(ID)) %>%
